---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Call Rust code from R using the 'extendr' crate

<!-- badges: start -->
[![R build
status](https://github.com/clauswilke/rextendr/workflows/R-CMD-check/badge.svg)](https://github.com/clauswilke/rextendr/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/rextendr)](https://CRAN.R-project.org/package=rextendr)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

## Installation

To install the package, run:
```r
remotes::install_github("clauswilke/rextendr")
```

Note that this will install the package but does not guarantee that the package can do anything useful. You will also need to set up a working Rust toolchain, including libclan/llvm-config support to run [bindgen](https://rust-lang.github.io/rust-bindgen/). See the [installation instructions for libR-sys](https://github.com/extendr/libR-sys) for help. If you can successfully build libR-sys you're good.

## Usage

Basic use example:

```{r}
library(rextendr)

# some simple Rust code with two functions
rust_src <- "use extendr_api::*;

#[extendr]
fn hello() -> &'static str {
    \"Hello, this string was created by Rust.\"
}

#[extendr]
fn add(a: i64, b: i64) -> i64 {
    a + b
}
"

rust_source(code = rust_src, quiet = TRUE)

# call `hello()` function from R
hello()

# call `add()` function from R
add(14, 23)
```

