---
title: "Using Rust code in R Markdown documents"
author: "Claus O. Wilke"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{RMarkdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

## R Markdown

The rextendr package enables a new chunk type for knitr, `extendr`, which compiles and evaluates Rust code.

First, you need to load the rextendr package in a regular R chunk. You would normally do this in a hidden setup chunk at the top of your R Markdown document.
```{r}
library(rextendr)
```


Then, you can create a code chunk of type `extendr`. For example:
````markdown
`r ''````{extendr}
rprintln!("Hello from Rust!");

let x = 5;
let y = 7;
let z = x*y;
z.into()
```
````
This chunk generates the following output:
```{extendr}
rprintln!("Hello from Rust!");

let x = 5;
let y = 7;
let z = x*y;
z.into()
```

The call `z.into()` casts the Rust integer variable into type `Robj` and returns its value to R. Notice the lack of a `;` at the end of this line. You can also write code that doesn't return any result to R. In this case, the last line would end in a `;`, as in this example:
```{extendr}
let x = 5;
let y = 7;
let z = x*y;

rprintln!("{}*{} = {}", x, y, z);
```

